---
title: "Quant Assignment 2"
author: "Alex Cox"
date: "9/14/2020"
output: html_document
---

```{r load libraries, results = 'hide', message=FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
```

```{r read CSV, results = 'hide', message=FALSE, echo=FALSE}
person_data <- read_csv("militarycommute.csv")
```

# Part I: Continuous Variables

For this exercise, I will first calculate the following 4 properties...

-Sample mean

-Sample standard deviation

-Interquartile range

-95-percent confidence interval for the population mean



For the following 3 continuous variables in my dataset:

1) Travel Time to Work (JWMNP)

2) Age (AGEP)

3) Income (PINCP)


# 1) Travel Time to Work (JWMNP)

```{r JWMNP summary, results = 'hide', message=FALSE, echo=FALSE}
person_data <- read_csv("militarycommute.csv")
JWMNP_summary <- summary(person_data$JWMNP)
JWMNP_summary
quantile(person_data$JWMNP)
sd(person_data$JWMNP)
conf_int <- t.test(person_data$JWMNP)
conf_int
```

```{r JWMNP info}
JWMNP_summary <- summary(person_data$JWMNP)
JWMNP_summary
paste("The sample mean for the travel time to work is",
      prettyNum(mean(person_data$JWMNP), digits = 0), "minutes.")
paste("The sample standard deviation for the travel time to work is",
      prettyNum(sd(person_data$JWMNP), digits = 0), "minutes.")
paste("The sample interquartile range for the travel time to work is",
      quantile(person_data$JWMNP)["25%"], "to", quantile(person_data$JWMNP)["75%"], "minutes.")
pop_mean <- t.test(person_data$JWMNP)
pop_mean
paste("The 95-percent confidence interval for the population mean is",
      prettyNum(pop_mean$conf.int[1], digits = 0), "to",
      prettyNum(pop_mean$conf.int[2], digits = 0), "minutes.")
```
```{r JWMNP histogram}
hist <- ggplot(person_data, aes(x = person_data$JWMNP)) +
  geom_histogram() 
hist
```

# 2) Age (AGEP)

```{r AGEP info}
AGEP_summary <- summary(person_data$AGEP)
AGEP_summary
paste("The sample mean for age is",
      prettyNum(mean(person_data$AGEP), digits = 0))
paste("The sample standard deviation for age is",
      prettyNum(sd(person_data$AGEP), digits = 0))
paste("The sample interquartile range for age is",
      quantile(person_data$AGEP)["25%"], "to", quantile(person_data$AGEP)["75%"])
pop_mean <- t.test(person_data$AGEP)
pop_mean
paste("The 95-percent confidence interval for age is",
      prettyNum(pop_mean$conf.int[1], digits = 0), "to",
      prettyNum(pop_mean$conf.int[2], digits = 0))
```
```{r AGEP histogram}
hist <- ggplot(person_data, aes(x = person_data$AGEP)) +
  geom_histogram() 
hist
```


# 3) Income (PINCP)

```{r PINCP info}
PINCP_summary <- summary(person_data$PINCP)
PINCP_summary
paste("The sample mean for income is",
      prettyNum(mean(person_data$PINCP), digits = 000000))
paste("The sample standard deviation for income is",
      prettyNum(sd(person_data$PINCP), digits = 00000))
paste("The sample interquartile range for income is",
      quantile(person_data$PINCP)["25%"], "to", quantile(person_data$PINCP)["75%"])
pop_mean <- t.test(person_data$PINCP)
pop_mean
paste("The 95-percent confidence interval for income is",
      prettyNum(pop_mean$conf.int[1], digits = 000000), "to",
      prettyNum(pop_mean$conf.int[2], digits = 000000))
```
```{r PINCP histogram}
hist <- ggplot(person_data, aes(x = person_data$PINCP)) +
  geom_histogram() 
hist
```


# Part II: Categorical Variables

Lastly, I will calculate the 95-percent confidence interval for the 2 categorical variables in my dataset:

4) Military Service (MIL)

5) Means of Transportation to Work (JWTR)

# 4) Military Service (MIL): 

```{r MIL summary}
unique(person_data$MIL_label)
table(person_data$MIL_label)
table(person_data$MIL_label) / sum(table(person_data$MIL_label))
t.test(person_data$MIL_label == "N/A (less than 17 years old)")
t.test(person_data$MIL_label == "Never served in the military")
t.test(person_data$MIL_label == "Now on active duty")
t.test(person_data$MIL_label == "On active duty in the past, but not now")
t.test(person_data$MIL_label == "Only on active duty for training in Reserves/National Guard")
```

# 5) Means of Transportation to Work (JWTR)

```{r JWTR summary}
unique(person_data$JWTR_label)
table(person_data$JWTR_label)
table(person_data$JWTR_label) / sum(table(person_data$MIL_label))
t.test(person_data$JWTR_label == "Car, truck, or van")
t.test(person_data$JWTR_label == "Walked")
t.test(person_data$JWTR_label == "Other method")
t.test(person_data$JWTR_label == "Bicycle")
t.test(person_data$JWTR_label == "Bus or trolley bus")
t.test(person_data$JWTR_label == "Subway or elevated")
t.test(person_data$JWTR_label == "Motorcycle")
t.test(person_data$JWTR_label == "Taxicab")
t.test(person_data$JWTR_label == "Railroad")
t.test(person_data$JWTR_label == "Ferryboat")
t.test(person_data$JWTR_label == "Streetcar or trolley car (carro publico in Puerto Rico)")
```
